LaMachine: A meta-distribution for NLP software

### 2018-07-26 docs/lamachine_extabstract_clarin2018.tex
[paper] minor edits

Being an open-source NLP software distribution, LaMachine is constrained to Unix-like platforms; this primarily means
Linux, but also BSD and, with some restrictions, macOS. Cygwin\footnote{A unix environment on Windows} is not tested or
supported. This constaint stands in contrast with mobile platforms (Android/iOS/etc), native Windows/mac desktop software, or
certain interface types in general such as classical desktop GUI applications, all of which fall beyond the scope of
LaMachine. Software that is incorporated must 1) bear some relevance to NLP, 2) be under a recognised open-source
license, 3) be deposited in a public version controlled repository\footnote{e.g. Github, Gitlab, Bitbucket, provided the
repository is public} and 4) have
a release protocol (with semantic versioning) using the proper technology-specific channels.

LaMachine is a \emph{meta distribution} as it can be installed in various contexts. At its core, LaMachine consists of a
set of machine-parsable instructions on how to obtain, install and configure software. These are implemented using
Ansible\footnote{\url{https://www.ansible.com}}.  This is notably different from the more classical notion of Linux
distributions, which generally provide their own repositories with (often binary) software packages. LaMachine builds on
this already available infrastructure by taking these repositories as a given and only needs to know which
repositories to use.  Similarly, there are different programming-language-specific ecosystems providing their own
repositories, such as the Python Package Index\footnote{\url{https://pypi.org}} for Python,
CRAN\footnote{\url{https://cran.r-project.org/}} for R, CPAN\footnote{\url{https://www.cpan.org}} for Perl, Maven
Central\footnote{\url{https://search.maven.org}} for Java.  LaMachine again relies on those to pull and install
software from and never forks, archives, or modifies the software in any way. In doing so, we compel participating
software projects to adhere to well-established distribution standards and ensure the software is more sustainable
towards the future \cite{softwarequality}. Moreover, we ensure that LaMachine never becomes a prerequisite for the software but merely a
courtesy or convenience.

@@

With LaMachine we also attempt to harmonise the metadata of all installed software, by converting metadata from upstream repositories to
the CodeMeta standard\footnote{\url{https://codemeta.github.io/}, described in JSON-LD} \cite{codemeta,codemetar} where possible, or encouraging software developers to provide
their own. This in turn enables other tools to do proper service discovery and provenance logging.

@@

We used a virtual machine (VM) image of LaMachine\footnote{Installation of LaMachine begins with a single bootstrap command, see
\url{https://proycon.github.io/LaMachine}} and succeeded in running it as a 64-bits Linux-based VM inside another
VM with Windows 2012 in which we did have administrative rights but no internet access, i.e. the VM that was offered to
us for this project. In terms of hardware we ran on a
machine with 8 cores and  32GB internal memory available.
%using VirtualBox and Vagrnt on the Windows VM
Lamachine provided a fully functional research environment and we ran all our experiments within LaMachine. We interacted with LaMachine both through the command line, which offers a standard
shell and enables access to all lower-level tools and programming languages; or through the (offline) webbrowser to use the Jupyter Notebook environment\footnote{\url{https://jupyter.org/}}, which offers a more graphical scripting environment
for e.g. Python and R that has gained great popularity in the scientific community.

>>>>

Being an open-source NLP software distribution, LaMachine is constrained to Unix-like platforms; this primarily means
Linux, but also BSD and, with some restrictions, macOS. Cygwin\footnote{A unix environment on Windows} is not tested or
supported. This constraint stands in contrast with mobile platforms (Android/iOS/etc), native Windows/mac desktop
software, or certain interface types in general such as classical desktop GUI applications, all of which fall beyond the
scope of LaMachine. However, virtualisation technology enables deployment on a wider range of platforms, including
Windows.  All software that is incorporated in LaMachine must 1) bear some relevance to NLP, 2) be under a recognised
open-source license, 3) be deposited in a public version controlled repository\footnote{e.g. Github, Gitlab, Bitbucket,
provided the repository is public} and 4) have a release protocol (with semantic versioning) using the proper
technology-specific channels.

LaMachine is a \emph{meta distribution} as it can be installed in various contexts. At its core, LaMachine consists of a
set of machine-parsable instructions on how to obtain, build (e.g. compile from source), install and configure software.
These are implemented using Ansible\footnote{\url{https://www.ansible.com}}.  This is notably different from the more
classical notion of Linux distributions, which generally provide their own repositories with (often binary) software
packages. LaMachine builds on this already established infrastructure by taking these repositories as a given and only
needs to know which repositories to use.  Similarly, there are different programming-language-specific ecosystems
providing their own repositories, such as the Python Package Index\footnote{\url{https://pypi.org}} for Python,
CRAN\footnote{\url{https://cran.r-project.org/}} for R, CPAN\footnote{\url{https://www.cpan.org}} for Perl, Maven
Central\footnote{\url{https://search.maven.org}} for Java.  LaMachine again relies on those to pull and install software
from and never forks, archives, or modifies the software in any way. In doing so, we compel participating software
projects to adhere to well-established distribution standards and ensure the software is more sustainable towards the
future \cite{softwarequality}. Moreover, we ensure that LaMachine never becomes a prerequisite for the software but
merely a courtesy or convenience.

@@

LaMachine we also aims to harmonise the metadata of all installed software, by converting metadata from upstream repositories to
the CodeMeta standard\footnote{\url{https://codemeta.github.io/}, described in JSON-LD} \cite{codemeta,codemetar} where possible, or encouraging software developers to provide
their own. This in turn enables other tools to do proper service discovery and provenance logging.

@@

We used a virtual machine (VM) image of LaMachine. Installation of LaMachine always begins with a single bootstrap
command\footnote{see
\url{https://proycon.github.io/LaMachine}} and we opted to build a VM and succeeded in running this 64-bits Linux-based VM inside another
VM with Windows 2012 in which we did have administrative rights but no internet access, i.e. the VM that was offered to
us for this project. In terms of hardware we ran on a
machine with 8 cores and  32GB internal memory available.
%using VirtualBox and Vagrnt on the Windows VM
Lamachine provided a fully functional research environment and we ran all our experiments within LaMachine. We interacted with LaMachine both through the command line, which offers a standard
shell and enables access to all lower-level tools and programming languages; and through the (offline) webbrowser to use the Jupyter Notebook environment\footnote{\url{https://jupyter.org/}}, which offers a more graphical scripting environment
for e.g. Python and R that has gained great popularity in the scientific community. Moreover, LaMachine also offers a portal
website with access and overview of all locally installed tools, including web services and web applications.
%Dit was er nog niet in de use-case formeel gezien
